### 2019-04-17 paper/response.txt
more

We are very grateful for the helpful comments that all reviews make, and are
especially gladened by the expressions of appreciation:

  * "extremely well-written"
  * "going significantly beyond [...] previous work"
  * "exemplary in its precision and rigor" 
  * " I enjoyed reading this paper"
  * ...

We will do our best to address all comments in the final version, but focus
here only on the main issues and questions raised.

- Review B:

  About Dom not having a bind:

  There are two key ideas that make the emulation of local state by means
  of local state work:

    1. We impose a discipline that all uses of put are in a state-restoring
       fashion; no `wild' puts are allowed.

       => Without a bind, a carefully constructed disciplined Dom value cannot
          be extended in a sequential (i.e., bind-) fashion with a wild put.

    2. The state at the end of a branch is transferred to (the beginning of)
       the next branch.

       => Without a bind, we know where a branch is at its end and can build in
          the transfer of the state to the next branch. With a bind, we would
	  not know when we are at the end of the branch and would have to defer
          the transfer, leaving room for the possibility that an entirely different
          state value is passed to the next branch.:w

  We will articulate this better. You make a good point here about moving the concrete Dom 
  from the appendix to the main text. Then we will be able to constrast with ListT-done
  right applied to the state monad, which suffers from the above two problems.

- Review C:

  Novelty wrt "Just do It":
  TODO

  Jeremy Gibbons, one of the authors of "Just do It" acknolwedges our contributions as follows:

> The Just Do It paper deals only with the backtrackable combination of
> nondeterminism and state, where composition distributes rightwards over failure
> (ie failure is a right zero of composition) and over choice - your laws
> (15,16). We don’t deal with persistent state, your law (18). We didn’t know
> then what the right axiomatization was for persistent state, nor a suitable
> model for it - we confidently asserted (S7) that s->([a],s) is an alternative
> model of nondeterminism and state, but in fact I believe this is not even a
> monad. Your paper contributes a useful axiomatization, experimental validation
> of those axioms through a variant of N Queens that doesn’t copy the state, a
> model, and a Coq verification of that model.

  The use of Haskell:
  We feel that the use of Haskell for programming and reasoning with monads is justified given
  the extensive existing body of literature that does just that. Nevertheless,
  we appreciate your pointing out that the accessibility of the paper will be improved when we 
  add more explanation of concepts like hylomorphisms and scanl.

  Benefits of Monads:
  A big benefit of monads is that it provides a pureful functional model of
  effects. This has become an actual practice in Haskell and is a good model for
  what happens in languages with algebraic effects & handlers. Yet, we should
  emphasize that the focus of our paper is reasoning about monadic programs, not
  to justify the use of monads. We are not aware of any other formalism used
  to tackle the same problem for other effect models. Even if it would exist,
  we believe that our work is still relevant because it directly applies to 
  the above settings where monads are used.

  (Non-)Modularity:
  There is modularity and non-modularity at different levels, which should not
  be mixed up.

    As we require a particular interaction between state and non-determinism,
    namely the local state semantics, by definition we do not want to reason about
    the two effects in isolation.

    Reasoning about local state happens modularly, and not for a whole program 
    at once. We use the local state laws for that.

    Enabling this local reasoning about local state, when implemented by means of
    global state, requires a global discipline of modifying state only in a 
    restoring-upon-backtracking fashion. We carry out this proof once, which involves
    reasoning about all disciplined whole programs, so that we can then reason
    locally about all specific programs.

    (Algebraic) Effect Handlers Approach:
    "in the middle of the paper it is revealed that the reader should have read
     the paper interpreting certain things as syntax and certain as semantics. In
     addition the meaning (semantics) of = is described in vague terms informally in
     a cryptic note reference to event handlers"

    We do not intend for the the reader to have read the first part of the
    paper in this way. On the contratry, like the "Just do It" paper, we advocate
    high-level axiomatic reasoning based on algebraic laws to understand programs. 
    However, as an auxiliary proof device, briefly sketched in the note on p.8, and
    elaborated in more detail in Sections 6-7 (and in full detail in our Coq 
    mechanisation), we can reason about monadic programs by separating their
    synax and semantics. Specific notions of equality are given on lines 942 and
    979.

    The second part of the paper presents a proof technique,
    where 
    (Note that we never mention event handlers, which are quite different from effect handlers.)

    TODO

- Review D:

   TODO

- A1:

  We are grateful for a better reference of Prolog's 4-port model.